---
title: "Over-simplistic approach to create a R data package"
description: |
  This is a note to myself! 
author: Bhabishya Neupane
date: 2022-03-28
preview: https://miro.medium.com/max/1400/1*dhGDw_mJ2yt7i9VwehDsiQ.png
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 4
twitter:
  site: "@_bhabishya_"
  creator: "@_bhabishya_"
categories:
  - usethis
  - pkgdown
  - tutorial
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# How to go about it?

Again, this will be more of a note to myself. 

1. Create a new github repo. 
2. Create a new RStudio project. 
3. Clone the repo and make a version control project.

4. Get all the essentials by running the following piece of code:

```{r eval = FALSE}
usethis::create_package(path = "give your path") # Make sure to use `getwd()` full file path. 
```

5. If we have raw data, then run:

```{r eval = FALSE}
usethis::use_data_raw()
```

After doing so, you'll see a `data-raw` folder in your project directory.

6. In `data-raw` folder, we'll have `DATASET.R` file where we can clean the script. 

7. On that very script, at the very bottom make sure to update the cleaned data. 

```{r eval = FALSE}
usethis::use_data(name_of_your_cleaned_data, overwrite = TRUE)
```

8. Do a quick check using:

```{r eval = FALSE}
devtools::check(document = FALSE)
```

9. You'll probably get some warnings and messages, read them carefully. 

10. Document the datasets (example from my own package):

```{r eval=FALSE}
usethis::use_r("name_of_your_cleaned_dataset")
```

```{r eval = FALSE}
#' Dairy Foods Information on all dairy plants in the United States 
#'
#' Dairy industry directory of dairy processors that supply milk, milk-based beverages, cheese, yogurt, ice cream, and other dairy foods and beverages
#'
#' @format A tibble with 1,983 rows and 35 variables:
#' \describe{
#'   \item{company_name}{chr Name of the Dairy Organization}
#'   \item{company_type}{chr Products the Dairy Organization specialized in} 
#'   \item{primary_product1}{chr denoting first primary product}
#'   \item{primary_product2}{chr denoting second primary product}
#'   \item{primary_product3}{chr denoting third primary product}
#'   \item{primary_product4}{chr denoting fourth primary product}
#'   \item{primary_product5}{chr denoting fifth primary product}
#'   \item{primary_product6}{chr denoting sixth primary product}
#'   \item{primary_product7}{chr denoting seventh primary product}
#'   \item{primary_product8}{chr denoting eighth primary product}
#'   \item{primary_product9}{chr denoting ninth primary product}
#'   \item{primary_product10}{chr denoting tenth primary product}
#'   \item{primary_product11}{chr denoting eleventh primary product}
#'   \item{primary_product12}{chr denoting twelfth primary product}
#'   \item{primary_product13}{chr denoting thirteenth primary product}
#'   \item{primary_product14}{chr denoting fourteenth primary product}
#'   \item{primary_product15}{chr denoting fifteenth primary product}
#'   \item{primary_product16}{chr denoting sixteenth primary product}
#'   \item{primary_product17}{chr denoting seventeenth primary product}
#'   \item{primary_product18}{chr denoting eighteenth primary product}
#'   \item{primary_product19}{chr denoting nineteenth primary product}
#'   \item{primary_product20}{chr denoting twentieth primary product}
#'   \item{primary_product21}{chr denoting twenty-first primary product}
#'   \item{primary_product22}{chr denoting twenty-second primary product}
#'   \item{primary_product23}{chr denoting twenty-third primary product}
#'   \item{primary_product24}{chr denoting twenty-fourth primary product}
#'   \item{primary_product25}{chr denoting twenty-fifth primary product}
#'   \item{primary_product26}{chr denoting twenty-sixth primary product}
#'   \item{primary_product27}{chr denoting twenty-seventh primary product}
#'   \item{primary_product28}{chr denoting twenty-eighth primary product}
#'   \item{primary_product29}{chr denoting twenty-ninth primary product}
#'   \item{primary_product30}{chr denoting thirty primary product}
#'   \item{primary_product31}{chr denoting thirty-first primary product}
#'   \item{primary_product32}{chr denoting thirty-second primary product}
#'   \item{primary_product33}{chr denoting thirty-third primary product}
#' }
#' @source <https://www.dairyfoods.com/directories>

"dairy_plant_data"
```

You can copy this whole thing and modify as required. 

11. Now, we need to document this!

```{r eval=FALSE}
devtools::document()
```

call the following once you're done documenting:

```{r}
?dairyplantdata #gives you the documentation
```



